Catecholamines and Hypoglycemia

### 2024-07-31 Christine Chan/EnVision CF/catecholamines_analysis.qmd
catecholamines tables done

```{r data cleaning}
# Import cleaned data
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# For this analysis, we only want those with catecholamines 
df = df %>% group_by(study_id,Date) %>%
  mutate(catecholamines = any(!is.na(Epinephrine)|!is.na(Norepinephrine))) %>%
  filter(catecholamines) %>% ungroup()
```

@@

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2,Hypo60,Hypo70)
  ) %>%
  select(Hypo70,age_visit:cftr_mutation_2,Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 70}
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model70
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo70)
```

@@

```{r}
#| label: tbl-60
#| tbl-cap: BG < 60 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 60}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model60
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo60)
```

# Hypoglycemia defined as <= 50 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-50
#| tbl-cap: BG <= 50 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 50}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model50
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo50)
```

>>>>

```{r data cleaning}
# Import cleaned data
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# For this analysis, we only want those with catecholamines
df <- df %>%
  group_by(study_id, Date) %>%
  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
  filter(catecholamines) %>%
  ungroup()
```

@@

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
#| warning: false
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2, Hypo70)
  ) %>%
  select(Hypo70, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

@@

```{r}
#| label: tbl-60
#| tbl-cap: BG < 60 at Any Point
#| warning: false
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2, Hypo60)
  ) %>%
  select(Hypo60, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 70} -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model70 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo70) -->
<!-- ``` -->

<!-- # Hypoglycemia defined as < 60 mg/dL at any point -->

<!-- ## Demographics -->

<!-- ```{r} -->
<!-- #| label: tbl-60 -->
<!-- #| tbl-cap: BG < 60 at Any Point -->
<!-- df %>% -->
<!--   select( -->
<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
<!--     all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno, -->
<!--     iAUC30gluc:iAUC180ins, homa_ir -->
<!--   ) %>% -->
<!--   tbl_summary(by = Hypo60, missing_text = "Missing") %>% -->
<!--   add_overall() %>% -->
<!--   add_p() %>% -->
<!--   separate_p_footnotes() -->
<!-- ``` -->

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 60} -->
<!-- # Long OGTTs -->
<!-- long_df <- df %>% -->
<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
<!--     names_to = "Time", values_to = "Glucose", -->
<!--     names_pattern = "timepoint_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model60 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo60) -->
<!-- ``` -->

<!-- # Hypoglycemia defined as <= 50 mg/dL at any point -->

<!-- ## Participant demographics -->

<!-- ```{r} -->
<!-- #| label: tbl-50 -->
<!-- #| tbl-cap: BG <= 50 at Any Point -->
<!-- df %>% -->
<!--   select( -->
<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
<!--     all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno, -->
<!--     iAUC30gluc:iAUC180ins, homa_ir -->
<!--   ) %>% -->
<!--   tbl_summary(by = Hypo50, missing_text = "Missing") %>% -->
<!--   add_overall() %>% -->
<!--   add_p() %>% -->
<!--   separate_p_footnotes() -->
<!-- ``` -->

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 50} -->
<!-- # Long OGTTs -->
<!-- long_df <- df %>% -->
<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
<!--     names_to = "Time", values_to = "Glucose", -->
<!--     names_pattern = "timepoint_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model50 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo50) -->
<!-- ``` -->
